---
title: "The `hsdar` package in R: Dealing with hyperspectral images"
subtitle: "Preface"
author: "Patrick Schratz"
date: "`r format(Sys.time(), '%a %d %b %Y')`"
output:
  reportMD::multi_document:
    toc: true
    fig_caption: true
    highlight: dracula
    depends:
      part1: 01_preprocessing.Rmd
      part2: 02_vegetation-indices.Rmd
      part3: 03_plot-speclibs.Rmd
      part4: 04_masking-and-filtering.Rmd
      part5: 05_continuum-removal.Rmd
    use_namespace: true
    css: R.css
params:
  version: !r 
---

```{r 00-hsdar-1, include=FALSE, cache=FALSE}
## load additional packages in this chunk
library(reportMD)
library(knitr)
library(pander)
library(ggplot2)
library(plotly)
```

```{r setup, include=FALSE, cache=FALSE}
## This chunk should contain global configuration commands.
## Use this to set knitr options and related things. Everything
## in this chunk will be included in an appendix to document the
## configuration used.

## Pander options
panderOptions("digits", 3)
panderOptions("table.split.table", 160)
```

```{r 00-hsdar-2, include=FALSE}
## Custom functions used in the analysis should go into this chunk.
## They will be listed in their own section of the appendix.

```

# Structure of this document

Some notes on the structure of this hands-on tutorial document using the `hsdar` package for hyperspectral data processing:  
All processing steps are split up into child documents. You can find these in the [Appendix] under the tab 'Related Documents'.

This file is a standalone html file which contains all code and images and can be shared without the .Rmd files. 
Information about 'Custom Functions', 'Document Configuration', 'R Session Info' and 'Child Documents' are provided in the [Appendix] for each single file.

This document uses the *Multi-part report* template from the [*reportMD*](https://github.com/humburg/reportmd) package, which is designed
to produce a series of interconnected HTML documents.

When you click the **Knit** button of the main Rmd file (`hsdar.Rmd`), all child documents will be generated including both figures/tables as well as the output of any embedded R code chunks within the document. 

The content of this compilation of documents borrows heavily from the vignette of the `hsdar` package (`vignette("hsdar-intro", package = "hsdar")`). 

```{r 00-hsdar-3, child='rmd/appendix.Rmd'}
```

