---
title: "Part 1: Data investigation & preprocessing"
author: "Patrick Schratz"
date: "`r format(Sys.time(), '%a %d %b %Y')`"
output: 
  html_document:
    theme: paper
    highlight: haddock 
    # code_folding: show
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE, cache=FALSE}
## knitr options
#knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(comment = "")
```

# Loading required packages

This is the first part of an hands-on tutorial using the `hsdar` package for hyperspectral data processing.  
At the beginning, we will load all necessary packages for this analysis. 

```{r 01-preprocessing-1, message=FALSE}
library("raster")
library("hsdar")
library("here")
library("rasterVis")
library("mapview")
library("magrittr")
library("stars")
```

# Reading data and creating a 'RasterBrick'

The hyperspectral image file for this analysis is shared with you in the MSCJ LIFE Slack group.  
Please download it and set the correct path.

The first step is to list all files within your working directory. This is the usual step when dealing with multiple files. 
Here, the file is stored in a subdirectory of this .Rmd file called 'data'. We list all files with a specific ending using the full path. 

```{r 01-preprocessing-2 }
# list files
file <- here("data") %>% 
  list.files(pattern = ".tif$", full.names = TRUE)
```

Next, we will read in these files as 'raster bricks' into R. 

```{r 01-preprocessing-3 }
raster <- raster::brick(file[1])
raster
```

Our image has 126 bands (nlayers), a spatial resolution of 1 meters and an UTM reference system (EPSG: 32630). 

# Visualization {.tabset .tabset-fade}

Let's take a quick look at the image. Band 100 is randomly chosen. 
Various options exists meanwhile in R for plotting spatial grid data:

- _rasterVis_
- _mapview_
- _stars_
- _ggplot2_ (via stars)

## rasterVis

```{r 01-preprocessing-4 }
levelplot(raster[[97:100]], margin = FALSE, pretty = TRUE)
```

## stars

```{r 01-preprocessing-5 }
img = stars::read_stars("data/laukiz1.tif")
plot(img[,,,97:100])
```

## mapview

```{r 01-preprocessing-6 }
mapview(raster[[97:100]], na.color = "transparent", map.types = "Esri.WorldImagery")
```

_mapview_ also supports a grid splitting of individual bands:

```{r 01-preprocessing-7}
m1 = mapview(raster[[97]], na.color = "transparent", map.types = "Esri.WorldImagery")
m2 = mapview(raster[[98]], na.color = "transparent", map.types = "Esri.WorldImagery")
m3 = mapview(raster[[99]], na.color = "transparent", map.types = "Esri.WorldImagery")
m4 = mapview(raster[[100]], na.color = "transparent", map.types = "Esri.WorldImagery")

sync(m1, m2, m3, m4)
```


## ggplot2

```{r 01-preprocessing-8 }
ggplot() +
  geom_stars(data = img[,,,97:100]) +
  scale_fill_viridis_c() +
  ggthemes::theme_map()
```

# Transformation into 'Speclib' or 'HyperSpecRaster' 

Until now, we 'only' have a raster file of class `RasterBrick`. 
While there would be a lot of options for further analysis with this `RasterBrick` object, `hsdar` requires its own class(es) for further analysis.
There are two options: 
- For most `hsdar` functions an object of class `speclib` is required. 
- If you have large raster files and want to perform high-performance processing using the _raster_ package, class `HyperSpecRaster` is needed. 

The first step is to merge the wavelength information with the `RasterBrick` file.
There is no way around creating a vector by hand that stores the band information.

(This information was extracted from the metadata information of this image file.)

```{r 01-preprocessing-9 }
wavelength <- c(404.08, 408.5, 412.92, 417.36, 421.81, 426.27, 430.73, 435.20, 
                439.69, 444.18, 448.68, 453.18, 457.69, 462.22, 466.75, 471.29,
                475.83, 480.39, 484.95, 489.52, 494.09, 498.68, 503.26, 507.86,
                512.47, 517.08, 521.70, 526.32, 530.95, 535.58, 540.23, 544.88,
                549.54, 554.20, 558.86, 563.54, 568.22, 572.90, 577.60, 582.29,
                586.99, 591.70, 596.41, 601.13, 605.85, 610.58, 615.31, 620.05,
                624.79, 629.54, 634.29, 639.04, 643.80, 648.56, 653.33, 658.10,
                662.88, 667.66, 672.44, 677.23, 682.02, 686.81, 691.60, 696.40,
                701.21, 706.01, 710.82, 715.64, 720.45, 725.27, 730.09, 734.91,
                739.73, 744.56, 749.39, 754.22, 759.05, 763.89, 768.72, 773.56, 
                778.40, 783.24, 788.08, 792.93, 797.77, 802.62, 807.47, 812.32,
                817.17, 822.02, 826.87, 831.72, 836.57, 841.42, 846.28, 851.13,
                855.98, 860.83, 865.69, 870.54, 875.39, 880.24, 885.09, 889.94,
                894.79, 899.64, 904.49, 909.34, 914.18, 919.03, 923.87, 928.71,
                933.55, 938.39, 943.23, 948.07, 952.90, 957.73, 962.56, 967.39,
                972.22, 977.04, 981.87, 986.68, 991.50, 996.31)
```

Now is the point at which the `hsdar` package comes into play. 
Let's fusion our raster file and the wavelength information.

We use the function `HyperSpecRaster()` and `speclib()`. 
The first argument will be our `RasterBrick` file, the second one the vector with the wavelength information.

```{r 01-preprocessing-10 }
hyperspecs <- hsdar::HyperSpecRaster(raster, wavelength)
class(hyperspecs)
```

```{r 01-preprocessing-11 }
speclib <- hsdar::speclib(raster, wavelength)
class(speclib)
```

We now have two files: 

1. A raster file of class `HyperSpecRaster`. 
  With this, we can do calculations on every spectrum (equals every pixel here) of our dataset using the _hsdar_ package in combination with the _raster_ package. 
  However, plotting of spectra and other functions of the _hsdar_ package do not work with class `HyperSpecRaster`.

2. An object of class `speclib`. 
  With this, we can perform any task that the _hsdar_ package offers. 
  Hence, we will use this object for all further processing from this point onward. 
  
# Subsetting speclibs

By inspecting the file we notice that the first four bands are corrupt.
A spectra can be subsetted using `hsdar::mask()`.
In our case, we specify the start and end values of the affected wavelengths:

```{r 01-preprocessing-12}
mask(speclib) <- c(404, 418)
speclib
```

In the next document (`vegetation-indices.Rmd`), we will calculate different vegetation indices. 
